---
title: "pastexample"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

text

<br>

## State and regional Plots
##### Cows that have calved Inventory, 1980 - 2020


Notes on abbreviations: 
ctc = "Cows that Calved"
og = original
NCBA  = National Cattleman's Beef Assoc.
R1, R2, R3, etc.. = Region1, Region2, Region3, etc.. 

PROJECT GOAL: Create Regional Plots (7 plots based on NCBA regions 1:7) with state totals of "ctc" (cows that calved) over time. Create one summary line plot with one line per region incorder to compare regional totals.

Workspace: is on dropbox, the LMIC excel is the original data with regional and national totals and all dates.REMEMBER FOR REPRODUCIBILITY, CHANGE WORKSPACE (IF NECESSARY) TO MATCH LOCATION OF THE EXCEL FILE!!


```{r}
# warning
# messgaes 
# results 
# echo 


#############    Begin Code: set workspace, load libraries     #############

remove(list = ls()) # deletes all saved objects
setwd("/Users/hrf/Dropbox/NWFCS_Writing and Code Samples/Code/")


#Load Libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggrepel)
library(gridExtra)
library(cowplot)
library(ggpubr)


# loading the FULL Dataset --------------------------------------
datafull_og_ctc<-read_xlsx("LMIC_full_dataset_total.xlsx", sheet=4, colnames(TRUE)) 
# NOTE SHEET 4 IS COWS THAT HAVE CALVED!!! ## <-- Change Sheet number for each sector (steers, replacements, feeders, etc..)

#############    Cleaning     #############
# Select: States col and 1980 - 2021 -----------
datafull_ctc_80toPrs<-datafull_og_ctc %>% select('States',`1980`:`2020`)
# Reshaping: Data wide -> data long form (and rename columns) ------------------------------
#Melt data form 80s to present
data_melt_ctc <- melt(datafull_ctc_80toPrs)
# name columns (note, just incase i misspelled state in the xls)...
colnames(data_melt_ctc) <- c("State","variable", "value")
# add the region back in
data_melt_ctc$region <- datafull_og_ctc$region
# data needs label column (labeling just the last occurance)
data_melt_ctc$label <- NA
data_melt_ctc$label[which(data_melt_ctc$variable == last(data_melt_ctc$variable))] <- data_melt_ctc$State[which(data_melt_ctc$variable == last(data_melt_ctc$variable))]
# Filtering: Keep all states (get rid of "Totals") -----------------------------------
# remove the totals (because just looking at state level, not regional totals for this plot)... use with %notin%
`%notin%`<-Negate(`%in%`) # creating the 'notin' operator
data_melt_ctc <- data_melt_ctc %>% filter(State %notin% c("R1 Total", "R2 Total","R3 Total","R4 Total","R5 Total","R6 Total","R7 Total","US Total"))
# note to self, here is another alternative to the %notin% alternative: 

#knitr::kable(head(data_melt_ctc %>% filter(!grepl('Total', State ))))
reactable::reactable(data_melt_ctc %>% filter(!grepl('Total', State )))
```

```{r}

# warning
# messgaes 
# results 
# echo 

#############    Regional Plots (7 plots): ggplot     #############
# ggPLOT: REGION 1 COWS THAT CALVES -----------------------------------------------
#plot (more aesthetic adjustments below)
R1_ctc <- ggplot(data_melt_ctc[data_melt_ctc$region == '1',], 
                 aes(variable, value, col = State, group=factor(State))) + 
  geom_line(aes(color=factor(State))) + 
  labs(x = "Year", y = "Number of Stock (In Thousands)", 
       caption = "includes dairy", title = "Region 1 - Cows That Have Calved") +
  theme(plot.caption = element_text(color = "grey", face = "italic"),
        plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 60),
        plot.title = element_text(color = "black",
                                  face = "bold", size = 12, family = "Arial")) +
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE)+
  theme(legend.position = "none")


# Setting the background 
R1_ctc <- R1_ctc + theme(
  panel.background = element_rect(fill = "white", colour = "grey",
                                  size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.1, linetype = 'solid',
                                  colour = "grey"), 
  panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                  colour = "grey"))

```

setting the axis scales

```{r}
summary(data_melt_ctc[data_melt_ctc$region == '1',]) #use in deciding scale
R1_ctc <- R1_ctc + scale_y_continuous(breaks=seq(0,2000,100)) 
R1_ctc <- R1_ctc + scale_x_discrete(breaks=seq(1980,2021,5)) #
```

View the final plot

```{r}
R1_ctc
```



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# echo means it wont show your code, just the plot, 
# results, does the reverse, if results =false it won't show the output

plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
